
\chapter{A Dynamical Approach to Quantum Algorithms}
\label{chap:dynamics}
\index{Quantum Algorithms@\emph{Quantum Algorithms}!Dynamical Approach@\emph{Dynamical Approach}}%

%--------------------------------------------------------------------------------   
%--------------------------------------------------------------------------------   
\section{Grover}

\label{sec:sincos}

Grover's algorithm was introduced and discussed at great length in
chapter (\ref{chap:quantumAlgorithms}).  There, it was mentioned that
Grover's algorithm can be seen as a rotation of states in Hilbert
space.  This section will establish this result a little more rigorously 
and show how it is used in an analysis of the complexity of the algorithm.

Consider the discrete action of Grover's algorithm as described in the end of
section (\ref{sec:grover}).  This will describe a search for \emph{one},
$\ket\omega$,
out of a list of $N$ states.  At any given step in the process, the
quantum register is in the state
\begin{equation}
\ket{\Psi} = k\ket{\omega} + \sum_{j\in\omega_\perp}l\ket{j},
\end{equation}
for some $k$ and $l$, which, at say the $i$-th iteration, 
look like
\begin{equation}
\begin{split}
k_{i+1} =& \frac{N-2}{N}k_i + \frac{2(N-1)}{N}l_i\\
l_{i+1} =& -\frac{2}{N}k_i + \frac{N-2}{N}l_i.
\end{split}
\label{e:discsincos}
\end{equation}

\begin{prop}
The coefficients $k$ and $l$ in (\ref{e:discsincos}) obey
\begin{equation}
\begin{split}
k_i =& \sin\left[\left(2i+1\right)\theta\right]\\
l_i =& \frac{1}{\sqrt{N-1}}\cos\left[\left(2i+1\right)\theta\right]
\end{split}
\label{e:sincos}
\end{equation}
for $\sin\theta = \frac{1}{\sqrt{N}}$.
\end{prop}

\begin{proof}
(By induction)  
Starting at step zero with an equally weighted 
superposition of all states
\begin{equation}
\ket{\Psi_0} = \frac{1}{\sqrt{N}}\ket{\omega} 
                + \sum_{j\in\omega_\perp}\frac{1}{\sqrt{N}}\ket{j}
\end{equation}
implies $k_0 = l_0 = \frac{1}{\sqrt{N}}$.
Inserting $i=0$ into (\ref{e:sincos}), you get
\begin{equation}
k_0 = \sin\left(\theta\right) = \frac{1}{\sqrt{N}}
\end{equation}
and 
\begin{equation}
\begin{split}
l_0 =& \frac{1}{\sqrt{N-1}}\cos\left(\theta\right)\\
    =& \frac{1}{\sqrt{N-1}}\sqrt{ 1 - \sin^2\left(\theta\right) }\\
    =& \frac{1}{\sqrt{N-1}}\sqrt{ 1 - \frac{1}{N} }
    = \frac{1}{\sqrt{N}}.
\end{split}
\end{equation}
So, the proposition holds for $i=0$.

Now, consider the $(i+1)$--th iteration
\begin{equation}
\begin{split}
&\begin{cases}
k_{i+1} =& \sin\left[\left(2(i+1)+1\right)\theta\right]\\
l_{i+1} =& \frac{1}{\sqrt{N-1}}\cos\left[\left(2(i+1)+1\right)\theta\right]
\end{cases}\\
&\begin{cases}
k_{i+1} =& \sin\left[\left(2i+1\right)\theta+2\theta\right]\\
l_{i+1} =& \frac{1}{\sqrt{N-1}}\cos\left[\left(2i+1\right)+2\theta\right]
\end{cases}\\
&\begin{cases}
k_{i+1} =& \sin\left[\left(2i+1\right)\theta\right]
           \cos\left(2\theta\right)
          +\cos\left[\left(2i+1\right)\theta\right]
           \sin\left(2\theta\right) \\
l_{i+1} =& \frac{1}{\sqrt{N-1}}\left\lbrace
                \cos\left[\left(2i+1\right)\theta\right]
                \cos\left(2\theta\right)
               -\sin\left[\left(2i+1\right)\theta\right]
                \sin\left(2\theta\right) 
           \right\rbrace.
\end{cases}
\end{split}
\end{equation}
By the induction hypothesis, 
\begin{equation}
\begin{cases}
k_{i+1} =& k_i\cos\left(2\theta\right)
          +\sqrt{N-1}\;l_i\sin\left(2\theta\right) \\
l_{i+1} =& \frac{1}{\sqrt{N-1}}\left\lbrace
                \sqrt{N-1}\;l_i\cos\left(2\theta\right)
               - k_i\sin\left(2\theta\right) 
           \right\rbrace,
\end{cases}
\end{equation}
and since
\begin{equation}
\sin\left(\theta\right) = \frac{1}{\sqrt{N}}
\quad\Rightarrow\quad
\cos\left(2\theta\right) = \frac{N-2}{N}
\quad\text{and}\quad
\sin\left(2\theta\right) = \frac{2\sqrt{N-1}}{N},
\end{equation}
then
\begin{equation}
\begin{cases}
k_{i+1} =& k_i\frac{N-2}{N} + l_i\frac{2(N-1)}{N}\\
l_{i+1} =& l_i\frac{N-2}{N} -k_i\frac{2}{N}. 
\end{cases}
\end{equation}
Since this is (\ref{e:discsincos}) as expected, $j\Rightarrow j+1$.
\end{proof}

%--------------------------------------------------------------------------------   
\subsection{Grover is Optimal}

\mmm{add Zalka stuff here}
\mmm{add in multisearch stopping time here too (from Boyer et al)}

%--------------------------------------------------------------------------------   
\subsection{Grover is a Geodesic}

Section (\ref{ssec:sincos}) expressed Grover's algorithm in terms of
discrete recurrence relations.  In the algorithm, the coefficients of
the state
\begin{equation}
\ket{\Psi} = k\ket{\omega} + \sum_{j\in\omega_\perp}l\ket{j}
\end{equation}
evolve according to
\begin{equation}
\begin{split}
k_i =& \sin\left[\left(2i+1\right)\theta\right]\\
l_i =& \frac{1}{\sqrt{N-1}}\cos\left[\left(2i+1\right)\theta\right].
\end{split}
\end{equation}
This evolution can be viewed as a discrete path in state space
as shown in figure (\ref{fig:discreteGrover}).
\begin{figure}[h]
\begin{center}
\begin{picture}(200,150)
    \thicklines
    \qbezier(0,70)(40,150)(120,100)
    \qbezier(120,100)(150,80)(180,85)
    \qbezier(0,70)(-40,-50)(140,0)
    \qbezier(180,85)(270,90)(200,30)
    \qbezier(200,30)(170,10)(140,0)
    \put(20,20){$\ket{\Psi}$}
    \put(110,70){$\ket{\omega}$}
    \qbezier[15](25,35)(70,70)(175,60)
\end{picture}
\caption{Grover's algorithm is a discrete path in state space 
starting from an equally--weighted superposition $\ket\Psi$ and 
continuing on through the desired state $\ket\omega$.  Note that
the algorithm keeps going and must be stopped at the anticipated
running time (When the algorithm will be near the desired state).}
\label{fig:discreteGrover}
\end{center}
\end{figure}

Now, consider the continuous path on the space of states
given by
\begin{equation}
\gamma\colon \R\to\CP^N\colon t\mapsto 
\begin{pmatrix}
    \sin(t)\\
    \frac{1}{\sqrt{N-1}}\cos(t)\\
    \vdots\\
    \frac{1}{\sqrt{N-1}}\cos(t)
\end{pmatrix}.
\label{e:gamma}
\end{equation}
The discrete points generated by Grover's algorithm 
{\sl lie upon this path}.  This path can then be used
as a {\sl continuous} approximation to Grover's algorithm
(figure \ref{fig:contGrover}).
\begin{figure}[h]
\begin{center}
\begin{picture}(200,150)
    \thicklines
    \qbezier(0,70)(40,150)(120,100)
    \qbezier(120,100)(150,80)(180,85)
    \qbezier(0,70)(-40,-50)(140,0)
    \qbezier(180,85)(270,90)(200,30)
    \qbezier(200,30)(170,10)(140,0)
    \put(20,20){$\ket{\Psi}$}
    \put(110,70){$\ket{\omega}$}
    \qbezier(25,35)(70,70)(175,60)
\end{picture}
\caption{Grover's algorithm can be approximated by a continuous
path in state space.  Note that this path coincides with the large
database limit of the algorithm and is a geodesic of the Fubini--Study
metric.}
\label{fig:contGrover}
\end{center}
\end{figure}

\begin{prop}
The path $\gamma$, as defined in (\ref{e:gamma}),
is a geodesic of the Fubini--Study metric on $\CP^N$.
\end{prop}

\begin{proof}
%In homogeneous coordinates, $\gamma$ is seen to be
%a geodesic of the $(2N+1)$--sphere.
%Since $\CP^N$ can be written as 
%\begin{equation}
%S^1\to S^{2N+1}\to \CP^N,
%\end{equation}
%then the geodesic projects down onto a geodesic of the base.
%\mmm{grin}
In inhomogeneous coordinates, $\gamma$ becomes
\begin{equation}
z_i = \frac{1}{\sqrt{N-1}}\cot(t)
\end{equation}
for all $i$.  A little sweat will show that these satisfy 
the equations of motion (\ref{e:detailedGeodesics}) for a geodesic
of the Fubini--Study metric on $\CP^N$.
\end{proof}


%\begin{prop}
%$\gamma$ coincides with the large database
%limit of Grover's algorithm.
%\end{prop}
%
%\begin{proof}
%Duh!
%\end{proof}
Note that $\gamma$ coincides with the large database
limit of Grover's algorithm.  This can easily be seen from
the large $N$ limit of (\ref{e:sincos}).

%--------------------------------------------------------------------------------   
\subsection{Is there a Hamiltonian?}

The diffusion transformation in Grover's algorithm,
as defined in section (\ref{ssec:diffusion}), looks like
\begin{equation}
{\bf U_s} =
\begin{pmatrix}
    \frac{2-N}{N} & \frac{2}{N} & \cdots & \frac{2}{N} \\
    \frac{2}{N}   & \frac{2-N}{N} & &  \\
    \vdots &  & \ddots & \vdots \\
    \frac{2}{N}   & \cdots && \frac{2-N}{N} 
\end{pmatrix}.
\end{equation}
With a little work, this can be rewritten as
\begin{equation}
{\bf U_s} =
- \exp \left\lbrace  i\frac{\pi}{N}
                \begin{pmatrix}
                    1      &        & \cdots &      & 1 \\
                    \\
                    \vdots &        & \ddots &      & \vdots \\
                    \\
                    1      &        & \cdots &      & 1
                \end{pmatrix}
\right\rbrace.
\end{equation}
For the purpose of argument, assume that the location (index) of
the target state is known ahead of time.  In this case, the
phase shift in Grover's algorithm (\ref{e:oracle}) can be written as
\begin{equation}
{\bf U_\omega} = 
\exp \left\lbrace  i\pi
                \begin{pmatrix}
                    0       &        &        &        &  \\
                            & \ddots &        &        &  \\
                            &        & 1      &        &  \\
                            &        &        & \ddots &  \\
                            &        &        &        & 0
                \end{pmatrix}
\right\rbrace,
\end{equation}
and so the entire Grover iteration can be written
\begin{equation}
\begin{split}
{\bf U}_{\text{Grover}} =& {\bf U}_s {\bf U}_\omega\\ 
=& - \exp \left\lbrace  i\frac{\pi}{N}
                \begin{pmatrix}
                    1      &        & \cdots &      & 1 \\
                    \\
                    \vdots &        & \ddots &      & \vdots \\
                    \\
                    1      &        & \cdots &      & 1
                \end{pmatrix}
\right\rbrace
\exp \left\lbrace  i\pi
                \begin{pmatrix}
                    0       &        &        &        &  \\
                            & \ddots &        &        &  \\
                            &        & 1      &        &  \\
                            &        &        & \ddots &  \\
                            &        &        &        & 0
                \end{pmatrix}
\right\rbrace.
\end{split}
\end{equation}
\mmm{Does this help find a Hamiltonian?}




%--------------------------------------------------------------------------------   
%--------------------------------------------------------------------------------   
\section{Dynamical Stability}
\label{sec:dynamicalStability}
\index{Dynamical Stability@\emph{Dynamical Stability}}%


The primary model used in this dissertation approximates 
Grover's algorithm as a geodesic of the Fubini--Study metric 
as discussed in the previous section.% \ref{sec:sincos}.

%--------------------------------------------------------------------------------   
\subsection{Linear Stability Analysis}

Inspection of the geodesic equations of motion 
(\ref{e:detailedGeodesics}) yields a fixed point about the 
origin.  No other fixed points are expected. \mmm{Why?}
The Jacobian of the evolution vanishes near the origin, and
so the equations of motion of the linearized problem become
the trivial geodesic equations for a flat metric.
This result is not particularly useful in characterizing how 
the algorithm will behave, at finite step size, in a noisy 
environment.  
A method of stability analysis
is therefore needed that does not rely upon a linearization
of the problem.  This gets quite difficult to do in general,
so a numerical approach is then warranted.

%--------------------------------------------------------------------------------   
\subsection{Numerical Model of Grover's Algorithm}

A database consisting of $N=2^n$ items, is represented (using
$n$ qubits) as an $N$--state system whose states are given as
points in $\CP^{N-1}$.  The evolution of the algorithm is then 
modelled by the numerical integration of the geodesic equations 
of motion (\ref{e:detailedGeodesics}) on this space.
Appropriate initial conditions are supplied to mimic the behaviour of 
the actual algorithm.  At this stage, the approach involves only pure 
states, and so the (unitary) evolution remains on the shell of pure 
states.
\mmm{show baseline results here?}
\mmm{uncertainties here?}
\mmm{pseudotime relates to number of Grover iterations}

\subsubsection{Simulation of Grover as a dynamical map}

The time evolution of a quantum mechanical system is most easily
described by unitary evolution.
Unfortunatley, physical systems do not exist in isolation.  There
are always interactions with environments or baths that cause
quantum mechanical systems to evolve by possibly more complicated
means than these simple unitary transformations.

A somewhat more general approach to quantum evolution is to 
consider dynamical evolution of a quantum mechanical system
described as simply a map, a \emph{dynamical map}\cite{Sudarshan:??}
of density matrices to density matrices.
Notice that an even more general approach is to look at maps of 
density matrices, where the image itself doesn't necessarily have to
be a physical density matrix.  It could be simply a subsystem
of a physical system, and hence it's trace could be less than
one.  For purposes of this dissertation, only dynamical maps that
take physical density matrices to physical density matrices are
considered.

\mmm{Phase space stuff or configuration space stuff}

The numerical simulation of Grover's algorithm described above
integrates the geodesic equations of motion.  At every time step,
this numerical integration can be viewed as simply a map that
takes density matrices to density matrices\dots a dynamical map.
\footnote{Really, this particular mapping maps derivatives as well, and
can be thought of as a mapping of the state in a quantum phase 
space.}

%--------------------------------------------------------------------------------   
\subsection{Noise}

Now, any noise introduced into the dynamical evolution of a 
quantum mechanical system can be described within the context 
of dynamical maps.  
No matter the source of the noise or whether the noise is 
Hamiltonian or non-Hamiltonian, this approach will encompass
it all.
\mmm{Include simple analytical perturbation if possible.  Use this to
compare with code}
\mmm{maybe example with simple problem as dynamical map...add noise
(physically motivated noise)}

Try two different models\dots
\begin{enumerate}
\item Run density matrices.  Take finite Grover iterations, then add
infinitesimal random Hermitian matrices.  This will model some general
noise in Grover(?). This will take pure to mixed states.
\item Represent mixed state density matrices as
\begin{equation}
\rho = \sum\lambda_i\rho_i
\end{equation}
and let the $\rho_i$ flow as pure states.  Add in noise to $\lambda$s
{\sl and} to the $\rho_i$ evolutions.
\end{enumerate}

\subsubsection{No Noise}
\subsubsection{...}
\subsubsection{Max Noise}



%--------------------------------------------------------------------------------   
%--------------------------------------------------------------------------------   
\section{Other Quantum Algorithms}

Shor?
